---
title: "Lab 6: Inference for numerical data"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    css: ../lab.css
---

```{r global_options, include=FALSE}
library(statsr)
library(tidyverse)
```

* * *

<div id="template">
**Your reproducible lab report:** Before you get started, download the R Markdown 
template for this lab. Remember all of your code and answers go in this document:
```{r load-template, eval=FALSE}
download.file("https://kelrenmor.github.io/STA101-002/labs/lab6/sta101_lab6.Rmd",
              destfile="sta101_lab6.Rmd")
```
</div>

## North Carolina births

In 2004, the state of North Carolina released a large data set containing 
information on births recorded in this state. This data set is useful to 
researchers studying the relation between habits and practices of expectant 
mothers and the birth of their children. We will work with a random sample of 
observations from this data set.

## R Packages

We will be using the `statsr` package that we used in Lab 5. We'll also be using the `tidyverse` library (which includes `ggplot2` for visualization and `dplyr` for data manipulation).

```{r statsr-load, eval=FALSE}
library(statsr)
library(tidyverse)
```

Remember, loading a package is something you need to do each
time you launch a new R session and in each of your R Markdown files that 
have code that relies on this package.

## Exploratory analysis

Load the `nc` data set into our workspace.

```{r load-data}
load(url("https://www.openintro.org/data/rda/ncbirths.rda"))
```

We have observations on 13 different variables, some categorical and some 
numerical. The meaning of each variable is as follows.

variable         | description
---------------- | ---------------------------------------------
`fage`           | father's age in years.
`mage`           | mother's age in years.
`mature`         | maturity status of mother.
`weeks`          | length of pregnancy in weeks.
`premie`         | whether the birth was classified as premature (premie) or full-term.
`visits`         | number of hospital visits during pregnancy.
`marital`        | whether mother is `married` or `not married` at birth.
`gained`         | weight gained by mother during pregnancy in pounds.
`weight`         | weight of the baby at birth in pounds.
`lowbirthweight` | whether baby was classified as low birthweight (`low`) or not (`not low`).
`gender`         | gender of the baby, `female` or `male`.
`habit`          | status of the mother as a `nonsmoker` or a `smoker`.
`whitemom`       | whether mom is `white` or `not white`.

1.  What are the cases in this data set? How many cases are there in our sample?

Remember that you can answer this question by viewing the data in the data viewer or
by using the following command:

```{r summary}
str(nc)
```

As you review the variable summaries, consider which variables are categorical 
and which are numerical. For numerical variables, are there outliers? If you 
aren't sure or want to take a closer look at the data, make a graph.

Consider the possible relationship between a mother's smoking habit and the 
weight of her baby. Plotting the data is a useful first step because it helps 
us quickly visualize trends, identify strong associations, and develop research
questions.

2.  Make a side-by-side boxplot of `habit` and `weight`. What does the plot 
highlight about the relationship between these two variables?

```{r boxplot}
ggplot(nc, aes(x=habit, y=weight)) + 
  geom_boxplot()
```

The box plots show how the medians of the two distributions compare, but we can
also compare the means of the distributions using the following to 
first group the data by the `habit` variable, and then calculate the mean
`weight` in these groups using the `mean` function.

```{r by-means}
nc %>%
  group_by(habit) %>%
  summarise(mean_weight = mean(weight))
```

There is an observed difference, but is this difference statistically 
significant? In order to answer this question we will conduct a hypothesis test.

## Inference

3.  Are all conditions necessary for inference satisfied? Comment on each. You can 
compute the group sizes using the same `group_by` command above but replacing `mean(weight)` 
with `n()`, as shown below.

```{r sample-sizes}
nc %>%
  group_by(habit) %>%
  summarise(sample_size = n())
```

Let $\mu_{non-smoker}$ denote the average weights of babies born to 
non-smoking mothers and $\mu_{smoker}$ denote the average weights of babies born to 
smoking mothers. The hypotheses for testing if the average weights of babies born to 
smoking and non-smoking mothers are different is as follows:

$$H_0: \mu_{non-smoker} = \mu_{smoker}$$
$$H_A: \mu_{non-smoker} \neq \mu_{smoker}$$

We can rearrange the above to make another equivalent way to write these hypotheses:

$$H_0: \mu_{non-smoker} - \mu_{smoker} = 0$$
$$H_A: \mu_{non-smoker} - \mu_{smoker} \neq 0$$

We revisit the `inference` function, that we will use for conducting
hypothesis tests and constructing confidence intervals. 

```{r inf-weight-habit-ht, tidy=FALSE}
inference(y = weight, x = habit, data = nc, statistic = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```

Let's pause for a moment to go through the arguments of this function. 
The first argument is `y`, which is the response variable that we are 
interested in: `weight`. The second argument is the explanatory variable, 
`x`, which is the variable that splits the data into two groups, smokers and 
non-smokers: `habit`. The third argument, `data`, is the data frame these
variables are stored in. Next is `statistic`, which is the sample statistic
we're using, or similarly, the population parameter we're estimating. Recall in Lab 5
we worked with "proportion". Next we decide on the `type` of inference 
we want: a hypothesis test (`"ht"`) or a confidence interval (`"ci"`). When performing a 
hypothesis test, we also need to supply the `null` value, which in this case is `0`, 
since the null hypothesis sets the two population means equal to each other. 
The `alternative` hypothesis can be `"less"`, `"greater"`, or `"twosided"`. 
Lastly, the `method` of inference can be `"theoretical"` or `"simulation"` based.

4.  Interpret the output of the above function call. What is the p-value? 
Do we reject the null hypothesis at the level $\alpha=0.05$?

5.  Now we change the `type` argument to `"ci"` to construct and record a confidence 
interval for the difference between the weights of babies born to nonsmoking and 
smoking mothers. Note that when doing a confidence interval arguments like
`null` and `alternative` are not useful (these are only relevant for hypothesis tests), 
so we removed them.

```{r inf-weight-habit-ci1, tidy=FALSE}
inference(y = weight, x = habit, data = nc, statistic = "mean", type = "ci", 
          method = "theoretical")
```

Interpret this interval in context of the data. Note that by 
default you'll get a 95% confidence interval. If you want to change the
confidence level, add a new argument (`conf_level`) which takes on a value
between 0 and 1.

By default the function reports an interval for ($\mu_{nonsmoker} - \mu_{smoker}$). 
We can easily change this order by using the `order` argument:

```{r inf-weight-habit-ci2, tidy=FALSE}
# Note the sign change on the CI (but same absolute values!)
inference(y = weight, x = habit, data = nc, statistic = "mean", type = "ci", 
          method = "theoretical", order = c("smoker","nonsmoker"))
```

* * *

## On your own

6. Calculate a 95% confidence interval for the average length of pregnancies 
(`weeks`) using the `inference` function and interpret it in context. 
Note that since you're doing inference  on a single population parameter, 
there is no explanatory variable, so you can 
omit the `x` variable from the function.

7. Calculate a new confidence interval for the same parameter at the 90% 
confidence level. You can change the confidence level by adding a new argument 
to the function: `conf_level = 0.90`. Comment on the width of this interval versus 
the one obtained in the the previous exercise.

8. Conduct a hypothesis test evaluating whether the average weight gained by 
younger mothers is different than the average weight gained by mature mothers.



<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0).
This lab was adapted for OpenIntro by Mine &Ccedil;etinkaya-Rundel from a lab 
written by the faculty and TAs of UCLA Statistics.
</div>