---
title: "Lab 6- Inference for numerical data"
author: "[Enter your name here]"
date: "[Enter date here]"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

```{r setup, include=FALSE}
# DO NOT ALTER CODE IN THIS CHUNK
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(statsr)
```

* * *

## Lab report

#### Load data:
```{r}
load(url("https://www.openintro.org/data/rda/ncbirths.rda"))
```

## Together:

#### Exercise 1: What are the cases in this data set? How many cases are there in our sample?

```{r summary}
str(nc)
```

#### Exercise 2: What does the plot highlight about the relationship between these two variables?

```{r boxplot}
ggplot(nc, aes(x=habit, y=weight)) + 
  geom_boxplot()
```

#### Exercise 3: Are all conditions necessary for inference satisfied? Comment on each.

```{r sample-sizes}
nc %>%
  group_by(habit) %>%
  summarise(sample_size = n())
```

#### Exercise 4: Interpret the output of the above function call (assume $\alpha=0.05$).

```{r inf-weight-habit-ht, tidy=FALSE}
inference(y = weight, x = habit, data = nc, statistic = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```

#### Exercise 5: Interpret this confidence interval in context of the data.

```{r inf-weight-habit-ci1, tidy=FALSE}
inference(y = weight, x = habit, data = nc, statistic = "mean", type = "ci", 
          method = "theoretical")
```

* * *

## On your own

#### Exercise 6: Insert code and answer question.



#### Exercise 7: Insert code and answer question.



#### Exercise 8: Insert code and answer question.


