---
title: "Lab 5: Inference for Categorical Data"
output:
  html_document:
    css: ../lab.css
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
---

<div id="template">
Before you begin the lab, download the lab report template. This template makes it very simple to include code and output in your write up from within RStudio as well as ensuring reproducibility of your results.
```{r load-template, eval = FALSE}
download.file("https://kelrenmor.github.io/STA101-002/labs/lab5/sta101_lab5.Rmd",
              destfile="sta101_lab5.Rmd")
```
</div>

In August of 2012, news outlets ranging from the [Washington
Post](http://www.washingtonpost.com/national/on-faith/poll-shows-atheism-on-the-rise-in-the-us/2012/08/13/90020fd6-e57d-11e1-9739-eef99c5fb285_story.html) to the [Huffington
Post](http://www.huffingtonpost.com/2012/08/14/atheism-rise-religiosity-decline-in-america_n_1777031.html)
ran a story about the rise of atheism in America. The source for the story was 
a poll that asked people, "Irrespective of whether you attend a place of 
worship or not, would you say you are a religious person, not a religious 
person or a convinced atheist?" This type of question, which asks people to 
classify themselves in one way or another, is common in polling and generates 
categorical data. In this lab we take a look at the atheism survey and explore 
what's at play when making inference about population proportions using 
categorical data.

## Set seed

Since we'll be using the `sample` function in this lab, which involves randomness, we set a seed
at the beginning of out document for reproducibility.

### Set seed:
```{r}
# Insert code for setting seed in this code chunk
```

## Load packages

In this lab, we will explore and visualize the data using the **tidyverse** suite of packages. We'll also use the **statsr** package for inference. Let's load the packages. Note that if running `library(statsr)` 
gives you a message saying `Error in library(statsr) : there is no package called ‘statsr’`
saying , then run `install.packages('statsr')` in your console.

```{r load-packages, message=FALSE}
library(tidyverse)
library(statsr)
```

## The survey

To access the press release for the poll, conducted by WIN-Gallup 
International, click on the following link:

*<https://sidmennt.is/wp-content/uploads/Gallup-International-um-tr%C3%BA-og-tr%C3%BAleysi-2012.pdf>*

Take a moment to review the report then address the following questions.

1.  In the first paragraph, several key findings are reported. Do these 
    percentages appear to be *sample statistics* (derived from the data 
    sample) or *population parameters*?

2.  The title of the report is "Global Index of Religiosity and Atheism". To
    generalize the report's findings to the global human population, what must 
    we assume about the sampling method? Does that seem like a reasonable 
    assumption?

## The data

Turn your attention to Table 6 (pages 14 and 15), which reports the
sample size and response percentages for all 57 countries. While this is
a useful format to summarize the data, we will base our analysis on the
original data set of individual responses to the survey. Load this data
set into R with the following command.

```{r head-data, eval=FALSE}
load(url("http://www.openintro.org/stat/data/atheism.RData"))
```

3.  What does each row of Table 6 correspond to? What does each row of 
    `atheism` correspond to?

To investigate the link between these two ways of organizing this data, take a 
look at the estimated proportion of atheists in the United States. Towards 
the bottom of Table 6, we see that this is 5%. We should be able to come to 
the same number using the `atheism` data.

4.  Using the command below, create a new dataframe called `us12` that contains
    only the rows in `atheism` associated with respondents to the 2012 survey 
    from the United States. Next, calculate the proportion of atheist 
responses. Does it agree with the percentage in Table 6?

```{r us-atheism}
# create a new dataframe (us12)
us12 <- atheism %>% filter(nationality == "United States" & year == "2012")

# calculate the proportion of atheists
us12 %>% summarize(prop_atheist = mean(response=='atheist'))
```

## Inference on proportions

As was hinted at in Exercise 1, Table 6 provides *statistics*, that is, 
calculations made from the sample of 51,927 people. What we'd like, though, is 
insight into the population *parameters*. You answer the question, "What 
proportion of people in your sample reported being atheists?" with a 
statistic; while the question "What proportion of people on earth would report 
being atheists" is answered with an estimate of the parameter.

The inferential tools for estimating population proportion are: the confidence interval and the 
hypothesis test.

5.  Write out the conditions for inference to construct a 95% confidence
    interval for the proportion of atheists in the United States in 2012.
    Are you confident all conditions are met?

If the conditions for inference are reasonable, we can calculate
the standard error and construct the interval. As you've seen in your PSs, you can do this by hand, 
but here in Lab we'll allow the `inference` function to do it for us.

```{r us-atheism-ci, tidy = FALSE}
statsr::inference(data = us12, y = response, success = "atheist", 
                  statistic = "proportion", type = "ci", method = "theoretical")
```

Note that since the goal is to construct an interval estimate for a 
proportion, it's necessary to specify what constitutes a "success", which here 
is a response of `"atheist"`.

Although formal confidence intervals and hypothesis tests don't show up in the 
report, suggestions of inference appear at the bottom of page 6: "In general, 
the error margin for surveys of this kind is $\pm$ 3-5% at 95% confidence".

6.  Based on the R output, what is the margin of error for the estimate of the 
    proportion of atheists in US in 2012?

## Success-failure condition

The textbook emphasizes that you must always check conditions before making 
inference. For inference on proportions, the sample proportion can be assumed 
to be nearly normal if it is based upon a random sample of independent 
observations and if both $np \geq 10$ and $n(1 - p) \geq 10$. This rule of 
thumb is easy enough to follow, but it makes one wonder: what's so special 
about the number 10?

The short answer is: nothing. You could argue that we would be fine with 9 or 
that we really should be using 11. What is the "best" value for such a rule of 
thumb is, at least to some degree, arbitrary. However, when $np$ and $n(1-p)$ 
reaches 10 the sampling distribution is sufficiently normal to use confidence 
intervals and hypothesis tests that are based on that approximation.

We can further investigate the interplay between $n$ and $p$ and the shape of the 
sampling distribution by using simulations. In the following code, we simulate the 
process of drawing 5000 samples of size 1040 from a population with a true 
atheist proportion of 0.1. For each of the 5000 samples we compute $\hat{p}$ 
and then plot a histogram to visualize their distribution.

```{r sim-np}
# Set p and n, make empty vector to hold p_hat results for 5000 simulations
p <- 0.1
n <- 1040
p_hats <- rep(NA, 5000)

# Run 5000 simulations
for(i in 1:5000){
  samp <- sample(c("atheist", "non_atheist"), n, replace = TRUE, prob = c(p, 1-p))
  p_hats[i] <- sum(samp == "atheist")/n
}

# Plot using base R
hist(p_hats, main="p = 0.1, n = 1040", xlim=c(0, 0.2))
```

Aside -- we've used base R plotting above, but we could do the same thing with ggplot:

```{r ggplot_args, eval=FALSE}
# Turn into data frame to use ggplot()
df <- data.frame(p_hats = p_hats)

# Make histogram of all p_hats from 5000 simulations
ggplot(df, aes(x=p_hats)) +      # tell ggplot what data frame and data to look at
  geom_histogram(bins=40) +      # tell ggplot what type of plot we want
  ggtitle("p = 0.1, n = 1040") + # add title
  xlim(0, 0.2)                  # set x limit
```

The commands in the simulation code build up the sampling distribution of $\hat{p}$ using a 
`for` loop. You can read the sampling procedure for the first line of 
code inside the `for` loop as, "take a sample of size $n$ with replacement 
from the choices of atheist and non-atheist with probabilities $p$ and $1 - p$,
respectively." The second line in the loop says, "calculate the proportion of 
atheists in this sample and record this value." The loop allows us to repeat 
this process 5,000 times to build a good representation of the sampling 
distribution.

7.  Describe the sampling distribution of sample proportions at $n = 1040$ and 
    $p = 0.1$. Be sure to note the center, spread, and shape.\
    *Hint:* Remember that R has functions such as `mean` to calculate summary
    statistics.

8. Replicate the above simulation three more times but with modified sample
    sizes and proportions: for $n = 400$ and $p = 0.1$, $n = 1040$ and 
    $p = 0.02$, and $n = 400$ and $p = 0.02$. Describe the three new sampling distributions. 
    Based on these limited plots, how does $n$ appear to affect the 
    distribution of $\hat{p}$? How does $p$ affect the sampling distribution of sample proportions?

* * *

## On your own
<div class="oyo">

Remember to use the `inference` function when finding CIs. As 
always, write out the hypotheses for any tests you conduct and outline the
status of the conditions for inference.

9. Calculate confidence intervals for the 
    proportion of atheists in 2012 in two other countries of your choice, and 
    report the associated margins of error as in questions 5 and 6. 
    Be sure to note whether the conditions for inference are met. 
    Specifically, create new data sets for each of the two countries first, 
    and then use these data sets in the `inference` function to construct the confidence intervals.

The question of atheism was asked by WIN-Gallup International in a similar 
survey that was conducted in 2005. (We assume here that sample sizes have 
remained the same.) Table 4 on page 12 of the report summarizes survey results 
from 2005 and 2012 for 39 countries.

10. Is there convincing evidence that Spain has seen a change in its 
    atheism index between 2005 and 2012?
    *Hint:* Create a new data set for respondents from Spain. Form
    confidence intervals for the true proportion of athiests in both
    years, and determine whether they overlap.

11. If in fact there has been no change in the atheism index in the countries 
    listed in Table 4, in how many of those countries would you expect to 
    detect a change (at a significance level of 0.05) simply by chance?\
    *Hint:* Look in the textbook index under Type 1 error.

12. What does the CLT imply about the sampling distribution of the sample proportion?
    What does the CLT imply about the distribution of the data?

</div>

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0). 
This lab was written for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel.
</div>