---
title: "Lab 7- Intro to regression"
author: "[Enter your name here]"
date: "[Enter date here]"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

```{r setup, include=FALSE}
# DO NOT ALTER CODE IN THIS CHUNK
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

* * *

## Lab report

#### Load data:
```{r}
load(url("https://kelrenmor.github.io/STA101-002/data/mlb11.Rdata"))
```

## Together:

#### Exercise 1:  What type of plot would you use to display the relationship between `runs` and one of the other numerical variables? Plot this relationship using the variable `at_bats` as the explanatory variable. Does the relationship look linear? If  you knew a team's `at_bats`, would you be comfortable using a linear model to predict the number of runs?

```{r}
### Plot the relationship between runs and at_bats
# [Insert plot here]
```

_Insert answers to questions here_

#### Exercise 2: Looking at your plot from the previous exercise, describe the relationship between these two variables. Make sure to discuss the form, direction, and strength of the relationship as well as any unusual observations.

```{r}
# See top of data set sorted by runs, 
# useful to see which team is the outlier
head(mlb11 %>% arrange(desc(runs)))
```

_Insert description here_

#### Exercise 3: Using `plot_ss`, choose a line that does a good job of minimizing the sum of squares. Run the function several times. What was the smallest sum of squares that you got? What if you intentionally choose a poorly fitting line -- how much larger does the sum of squares get?

_Insert answers to questions here_

#### Exercise 4: If a team manager saw the least squares regression line and not the actual data, how many runs would he or she predict for a team with 5,579 at-bats? Is this an overestimate or an underestimate for the Philadelphia Phillies, and by how much? In other words, what is the residual for this prediction?

```{r reg-with-line}
# Plot the data with the least squares regression line overlaid
ggplot(data = mlb11, aes(x = at_bats, y = runs)) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE)
```

```{r m1}
### Define linear model.
m1 <- lm(runs ~ at_bats, data = mlb11)

### Summarize linear model output.
summary(m1)
```

```{r}
# Look at Philadelphia Phillies data
mlb11 %>% filter(team=="Philadelphia Phillies")
```

_Insert answers to questions here_

#### Exercise 5: Is there any apparent pattern in the residuals plot? What does this indicate about the linearity of the relationship between runs and at-bats?

_Insert answers to questions here_

#### Exercise 6: Based on the histogram and the normal probability plot, does the nearly normal residuals condition appear to be met?

_Insert answers to questions here_

#### Exercise 7: Based on the residuals vs. fitted plot, does the constant variability condition appear to be met?

_Insert answers to questions here_

* * *

## On your own

#### Exercise 8: Choose another one of the seven traditional variables from `mlb11` besides `at_bats` (i.e., one of hits, home runs, batting average, strikeouts, stolen bases, and wins) that you think might be a good predictor of `runs`. Produce a scatterplot of the two variables and fit a linear model. At a glance, does there seem to be a linear relationship?

```{r}
### Define linear model.
# [Insert linear model code here]

### Create a scatterplot with the least squares line for the model laid on top.
# [Insert plot code here]
```

_Answer will depend on chosen variable._

#### Exercise 9: How does this relationship compare to the relationship between `runs` and `at_bats`? Use the R$^2$ values from the two model summaries to compare. Does your variable seem to predict `runs` better than `at_bats`? How can you tell?

```{r}
### Summarize linear model to get R^2.
# [Insert linear model summary code here]
```

_Answer will depend on chosen variable._

#### Exercise 10: Check the model diagnostics for the regression model using the three conditions we assessed above.

```{r}
### Condition 1: linearity.
# [Insert code to check condition 1 here]
```

_Answer will depend on chosen variable._

```{r}
### Condition 2: nearly normal residuals.
# [Insert code to check condition 2 here]
```

_Answer will depend on chosen variable._

```{r}
### Condition 3: constant variability.
# [Insert code to check condition 3 here if needed,
#  or use previous plots if that is sufficient]
```

_Answer will depend on chosen variable._

#### Exercise 11 (bonus -- if you do this correctly, you will be forgiven for up to 5 missed points on the remainder of the assignment, taking your total up to a max of 100 points): Now that you can summarize the linear relationship between two variables, investigate the relationships between `runs` and each of the other five  traditional variables, and also `runs` and each of the three newer variables (i.e. the statistics used by the author of *Moneyball*). Which variable seems to be the best predictor of `runs`? (Hint: figure out which has the highest R$^2$ by quickly running `lm()` using each possible explanatory variable and jotting down each of the resulting R$^2$ values.)

_The best predictor of `runs` seems to be..._

```{r}
### Define linear model for best predictor.
# [Insert linear model code here]

### Summarize linear model to get R^2 for best predictor.
# [Insert linear model summary code here]
```

_Support your conclusion by stating the R$^2$ value for this model._

_Using the limited (or not so limited) information you know about these baseball statistics, does your result make sense?_
